Early environmental conditions do not impact behavioural flexibility in an invasive and non-invasive lizard species

### 2024-03-16 ms.qmd
Updates

Pablo Recio-Santiago^1^,$\ddagger$ , Dalton L. Leibold^1^, Ondi Crino ^1,2^, Kris Wild^1,3^, Daniel W.A. Noble^1^

^1^ Division of Ecology and Evolution, Research School of Biology, The Australian National University, Canberra, ACT 2601, Australia &NewLine;
^2^ Flinder's University, Adelaide, SA 5001, Australia &NewLine;
^3^ Melbourne University, Melbourne, VIC 3000, Australia &NewLine;
$\ddagger$ Corresponding author: pablo.reciosantiago@anu.edu.au &NewLine;

## Abstract

@@

knitr::include_graphics("./Others/BEHFLEX_FIG_1.svg")

@@

>>>>

Pablo Recio-Santiago^1^,$\ddagger$ , Dalton C. Leibold^1^, Ondi Crino ^1,2^, Kris Wild^1,3^, Chris Friesen^4^, Basile Mauclaire^1,5^, Amelia Peardon ^1^, Daniel W.A. Noble^1^

^1^ Division of Ecology and Evolution, Research School of Biology, The Australian National University, Canberra, ACT 2601, Australia  
^2^ Flinder's University, Adelaide, SA 5001, Australia  
^3^ Melbourne University, Melbourne, VIC 3000, Australia  
^4^ University of Wollongong, Wollongong, NSW 2500, Australia  
^5^ Université de Lille, Lille 59000, France  
$\ddagger$ Corresponding author: pablo.reciosantiago@anu.edu.au

## Abstract
Behavioural flexibility is crucial for species adapting to new environments, specially during invasion events. However, the early environment experienced can profoundly impact animals' ability to behave flexibly. Aspects like parents' transmitted hormones or temperature are predicted to interact to shape behavioural flexibility. But, given the advantages associated to behaving flexibly during the invasion process, behavioural flexibility in invasive species is expected to be less susceptible to the impacts of early environment. Here, we investigated the interactive effects of prenatal corticosterone (CORT) and temperature on behavioural flexibility in two lizard species, _Lampropholis delicata_ and _L. guichenoti_, with distinct invasion success. We manipulated prenatal CORT levels and incubation temperature in a 2x2 factorial design, and then assessed behavioural flexibility through a reversal learning task. Contrary to our expectations, we found no significant differences in the reversal learning rate between treatments or species. Our results suggest that prenatal temperature and CORT may not influence reversal learning in these species and that other factors in _L. delicata_ biology may be more influential in invasion success than behavioural flexibility alone.

@@

knitr::include_graphics("./Others/BEHFLEX_FIG_1")

@@

## Ethics  
Experiments complied with Australian law and were approved by the Australian National University Animal Experimentation Ethics Committee (A2022_33).

## Data accessibility  
All data, data description, and R code are availabe in public repository https://github.com/Pablo-Recio/CORT_Temp_Behavioural_flexibility

## Declaration of AI use
We declare Chat GPT was used for questions related to coding and data analyses. All other parts of the manuscript were written by the authors and those parts were AI was used were checked and modified accordingly by the authors.

## Authors' contributions
P.R.: conceptualization, methodology, data collection, data curation, formal analysis, writing—original draft, writing—review and editing; D.C.L.: data collection, writing—review and editing; O.C.: conceptualization, methodology, writing—review and editing; K.W.: conceptualization, writing—review and editing; C.F.: conceptualization, methodology, writing—review and editing; B.M: data collection, writing—review and editing; A.P.: data collection, writing—review and editing; D.N.: conceptualization, methodology, funding acquisition, project administration, resources, supervision, writing—review and editing.  
All authors gave final approval for publication and agreed to be held accountable for the work performed therein.

## Conflict of interest declaration  
We declare we have no competing interests.

## Funding
This work was supported by a National Australian University fellowship (P.R.), the Australian Research Council (grant no. DP210101152) to D.N., and the ACT Herpetological Association grant to P.R.

## Acknowledgements
and two anonymous reviewers for their valuable feedback on the manuscript.
